% Shared document content — uses commands defined in lang-es.tex or lang-en.tex

\begin{titlepage}
	\centering
	\vspace{1cm}
	\vspace{1.5cm}
    \vfill
	{\huge\bfseries \docTitle\par}
	\vspace{2cm}
	{\Large Aarón Montoya Moraga\par}
    \vfill
	{\large \today\par}
\end{titlepage}

\newpage

\introText

En este documento se presentan los instrumentos de la serie Ciat-Lonbarde.

La primera vez que 

La primera vez que los vi en vivo fue en la tienda Control en Brooklyn, New York. En esa época estaba concentrándome en una mezcla de tocar instrumentos de cuerda con pedales, y a través de eso, 

Luego entré a estudiar un postgrado a MIT Media Lab, donde conocí a Will Freudenheim, 

Desde 2021 estoy de vuelta en Chile, y he reemplazando casi toda mi instrumentación en vivo por los instrumentos de Ciat-Lonbarde, y en ese proceso, he investigado mucho sobre ellos, y estos apuntes es mi forma de documentar y compartir este proceso.

\newpage

\chapter{Tetrax}

\introTetrax

\section{Entradas}

Fuente de poder

Batería de 9V DC

4 perillas

\section{Salidas}

Audio stereo

\chapter{Sidrax}

\introSidrax

\chapter{Quatrax}

\chapter{Deerhorn}


\chapter{Plumbutter}

\chapter{Rolzer}

\chapter{Cocoquantus}

\chapter{Clicker}

\chapter{Peterlin}

\chapter{Golden Section}

\chapter{Cafeteria\textunderscore Mono}

\chapter{Cafeteria\textunderscore Stereo}

\chapter{Cafe Quantum}


\chapter{Anexos}

\section{Textos de la web de Patch-Point}

\subsection{Sobre Tetrax}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-tetrax-organ}.

Wood 4 voice analog square wave based bass synthesizer.

The Tetrax Organ follows a lineage of electronic musical instruments based on touch. It was redesigned once, as Tetrazzi. It has four wooden bars that you press to express four oscillators within. Each oscillator has its own fine tuning slider, embedded right within its bar. Also, there is are two master pitch knobs, and two master chaos knobs. Each master chaos knob makes the instrument go from pure sawtooth tones to a nest of noise. It does this by gradually increasing modulation from one oscillator to the next in a circular pattern. The reason there are two of each knobs: the Tetrax is all about timbre modulation, by changing the upslope and downslope of triangle waves separately. So, for master pitch and chaos, there is a separate upwards and downwards knob. In general, Tetrax is for funky bass and complex tones, while Sidrax is for sweet melodies. Of course they all have noise available at any point!

There are 28 modulation points brought forth as banana jacks. They are arranged in four columns, representing the four bars. Their layout is shifted diagonally, to allow inputs in proximity to outputs. This is for touch-playability; you can make banana jack metal studs that can be squished by hand to gesture-patch the instrument. Inputs are cool colors: blue, green, and grey. Outputs are hot colors: red and orange. The grey jacks are parallel inputs with the master knobs: the two on the left are for master pitch, the right two are master chaos. The color allotment for each bar is as follows:

\begin{itemize}
	\item Two orange jacks, representing the press and release gestures on that bar in voltage
	\item One red jack, which is the pure triangle output of that oscillator
	\item Two green jacks, that are "glitch" inputs. They cause the boundary conditions of each oscillator to wire to the output of another. A modulation inspired by circuit bending
	\item One blue jack, an fm modulation input for that oscillator
\end{itemize}

The Tetrax organ has a stereo-mini output on the back. This is because each bar, outputs in different channels during the press/release cycle. In addition to the stereo-mini jack, it has an power switch, and 9volt battery snap.

Comes with universal PSU.

\newpage

\subsection{Sobre Sidrax}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-sidrax-organ}.

Triangle to chaos 7 voice polyphonic touching wood synthesizer organ

Text by Peter Blasser:

The Sidrax Organ follows a lineage of electronic musical instruments based on touch. It was redesigned twice, first as Sidrassi, then as Sidrazzi. It has seven wooden bars that you press to express seven oscillators within. Each oscillator has its own fine tuning slider, embedded right within its bar. Also, there is a master pitch knob, and a master chaos knob. The master chaos knob makes the instrument go from pure triangle tones to a nest of noise. It does this by gradually increasing modulation from one oscillator to the next in a circular pattern. Sidrax has enough bars to construct sweetest melodies from the triangle waves, whereas Tetrax is a simple four-bar entity, with complexity of tone for funky bass.

There are 44 modulation points brought forth as banana jacks. They are arranged in seven columns, representing the seven bars. Their layout is shifted diagonally, to allow inputs in proximity to outputs. This is for touch-playability; you can make banana jack metal studs that can be squished by hand to gesture-patch the instrument. Inputs are cool colors: blue, green, and grey. Outputs are hot colors: red and orange. The grey jacks are parallel inputs with the master knobs: the one on the left is for master pitch, the right one is master chaos. The color allotment for each bar is as follows:

\begin{itemize}
	\item Two orange jacks, representing the press and release gestures on that bar in voltage.
	\item One red jack, which is the pure triangle output of that oscillator.
	\item Two green jacks, that are "glitch" inputs. They cause the boundary conditions of each oscillator to wire to the output of another. A modulation inspired by circuit bending.
	\item One blue jack, an fm modulation input for that oscillator.
\end{itemize}

The Sidrax organ has a stereo-mini output on the back. This is because each bar, outputs in different channels during the press/release cycle. In addition to the stereo-mini jack, it has an power switch, and optional built-in batteries, or an external, 9volt battery snap.

Comes with universal 12v PSU

\newpage

\subsection{Sobre Quatrax}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-quatrax-organ}.

The Quatrax is a 14 voice polyphonic analog synthesizer, that accommodates a huge variety of sounds and expressive capabilities. It combines a new synth engine derived from two instruments: Sidrax and Tetrax, although it is based on principals of those two instruments, it is an entirely different sound engine of its own! 

\newpage

\subsection{Sobre Deerhorn}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-deerhorn}.

Text by Peter Blasser

The design of the Deerhorn circuit has evolved much since it was first created to answer Clara Rockmore's original complaint with the Theremin: "Can there be more than one pitch?". Lev Theremin responded that one would need more than two arms, but the Deerhorn makes this possible by extracting gestural information from the radio fields and using it to control pitch and envelope. The current unit contains three of these antenna-synthesizer modules, and adds two extra, highly useful modules known as intersections.

An intersection module simply samples one deerhorn's antenna voltage at impulses provided by another module. It is a sample and hold, triggered by changes in gestural direction. Since there are three modules, there are six permutations, doubled because each module reports changes in both directions; there are twelve new outputs, plus four special ones that sample in random steps. Each deerhorn module emits a different light— red, orange, and yellow— which also color-codes the intersection bananas to show which module is being sampled.

The Deerhorn modules are color coded. An orange banana reports its primary product: a control voltage representing the hand's distance from antenna. The impulses utilized in intersections, are also available as red bananas. These are like two complimentary square waves that flip when the hand changes direction.

The base tone of deerhorn is a simple two tone straight bellow. That's as your hand moves in towards the antenna and back out again. You must always tune the response first, so that the two leds cross each other at a comfortable range from the antenna. Fine tuning is done with the top knob.

The bottom cluster of controls deal with the synthesizer frequencies. The bottom knob sets the base pitch of both oscillators. The next two knobs are associated with the purple bananas: they are FM tuning knobs. The switch on the stub of the deerhorn orients the stereo position of the oscillators. In middle position, it turns them off (for the stereo auxiliary inputs). The white bananas output the same stereo sound which is output the main output, but in single tracks so it can be used as separate modulation sources.

For coarse tuning of the antenna, two trimmer capacitors lie beneath every deerhorn module. They are well-tuned upon shipping, but may always be accessed via a face-port with small slotted screwdriver.

\newpage

\subsection{Sobre Plumbutter}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-plumbutter-2}.

Drum and Drama Machine

Text by Peter Blasser

There is a new Plumbutter, which we will herein call "Plumbutter 2". It is very much similar to the old plumbutter, but in a new slim case. The sigils are engraved rather than printed. Still features the signature mulberry, walnut and sassafras woodage. Aux inputs and outputs remain. Besides housekeeping mods to clean the sound, there are two special new mods:

\begin{itemize}
	\item The stereo mixer section now has CV inputs. They act to in a stereo manner: a signal going up on "verso" makes the right side louder and the left side quieter, and the opposite for "inverso".
	\item Avdogs are now stereo: each one undulates in the left channel and the right channel, back and forth.
	\item The snare drum is now much louder
\end{itemize}


FOR THE REAL HEADSs. AUXILIARY INPUTS ARE ON THE FRONT, TO INTEGRATE IT WITH ANY SORT OF SYNTHESIS/PLAYBACK RIG. THE DEERHORN MODULE PROVIDES INSTANT SPATIAL CONTROL OF ANYTHING, VIA A SPECIAL PRINTED ANTENNA ON THE RIGHT SIDE OF THE INSTRUMENT. 

ONE THING YOU SHOULD KNOW IS THAT U NEED A LOT OF BANANA CORDS TO PLAY THIS ONE. PLEASE HAVE THEM READY FOR IT WHEN IT ARRIVES.

Here's a cool Hainbach video jamming on it!

\newpage

\subsection{Sobre Rolzer}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-rolzer}.

Ciat-Lonbarde Rolzer:

is a beat generator add-on for the Plumbutter, a drum and drama machine. The Plumbutter offers complete voltage control of all parameters, including its pulse-rolls, for generating beat tempo patterns. However, when voltage control is added to an androgynous circuit such as the rolz, it has the tends to roll-off the ultrasound chaos present in the original incarnation of this idea, the rolz-5. So here we have the Rolzer, a collection of geometries graded into tempi by virtue of their capacitors. It is for developing a relationship and searching for the oddest, most alive rhythms, not by twiddling knobs but by making connections.

According to the concepts outlined in the Rollz-5 papers, combinations of even and odd geometries, such as a square and a triangle, generate sophisticated beat patterns including chaos at ultrasound frequencies. An even roll maintains a steady pattern, and an odd roll tries to resolve its inherent paradox; connecting them blends these attributes over time and with "grain". The original paper circuits presented each solder-worker to construct her own idiosyncratic assemblage of geometries, with assortments of capacitors based on what was on at hand, or strict dogmata, or pure whimsey. When the concept evolved into a Laboratory Rollz, the Plumbutter discarded any idiosyncrasies for the sake of pure control and dialing in aesthetic by knob. However, the Tocante line offers a way to think about the materials of electronics- capacitors- that allows a revision of the original Rolz concept.

Here we can bring back the strong sense of idiosyncrasy only offered by non-knobbed geometries, without whimsey, by collaborating with the industrial value system of capacitors, known as "E6". Basically, a ratio of ten is broken logarithmically into six gradations, and these are mapped to a sense of tempo, associating Italian quality with Industrial quantity: presto, allegro, moderato, andante, adagio, lento, and grave become modules of 1, 1.5, 2.2, 3.3, 4.7, 6.8, and 10 microfarad values.

Now, each tempo module contains an assortment of even and odd rolls. From top to bottom: hexagon, triangle, square, pentagon, and hexagon again. There are two hexagons to offer that roll's plexus for connections. Before defining the colors, review the sandrode concept- androgynous nodes that neither input nor output, but perform both functions at different times. In the rolzer, they are simply transistor neurons, with three simple actions: discharge of a capacitor, building up to a threshold, and inverted pulsing another neuron. The sandrode function is at the junction of these three, forming both a sensor and a pulse generator. The rolzer modules' majority of connections are sandrodes, represented as brown nodes. In the language of plumbutter, brown is both a hot and a cool color, so it takes both those attributes, output and input. However, sometimes you just want an output, perhaps for voltage control of a manifesting gong. So, each roll has one orange jack that yields a square wave at the tempo of the roll.

Connections may be stacked, and since all the plumbutter and rollzer outputs have a current limiting resistor, anything can go into the brown bananas. However, note that the brown bananas have no current limiting resistor; this is essential to propagate the paradox wave effect. The impedance of a brown node feels like a diode junction to ground, plus a high impedance to positive. As of 2014, I have not received any complaints of broken browns, because most sensible synthesizers have some kind of current limiting resistor on their outputs. But make note that if you have a very old Serge, you will likely already know about its own smoking modules when outputs are crossed, and that old warning would apply when patching old Surge to a brown banana. In general, the browns are meant to be patched with each other, and connected to the green inputs of plumbutter for drum machine manifestation. Likewise, orange "should" go to blue or purple. And red and gray can come back into the browns here. But experimenting really can't hurt!

\newpage

\subsection{Sobre Cocoquantus}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/ciat-lonbarde-cocoquantus2}.

Sampler, Modulator, Delay and Looper Modular Synthesizer.

Cocoquantus fast facts by Peter Blasser:

\begin{itemize}
	\item it comes from a series of 8 bit digital delays, starting with the ambrazier, srine, and tranoe, then the cocolase, and most recently the cocostuber. all had two inputs and two outputs, for two separate digital delays (stereo)
	\item the digital delay being made from discrete silicon chips, it is "2012 ready:" with no flash memory 2 B erased by solar flares. as a discrete device it also offers unique possibilities for clock modulation of your samples.
	\item it incorporates something from each of the previous COCO incarnations, but also stands on its own as a new synthesizer.
	\item materials include: Walnut, Sassafras, Ash, Cherry, circuit board, transistors, capacitors, and resistors. It is RoHS compliant.
	\item the face of the instrument is an esoteric, Washington DC, spell, consisting of two obelisks and one pentagon in the middle.
	\item inside each obelisk is a little "counting computer" that samples your inputs and "loops" them.
	\item your inputs for each obelisk (COCO digital delay) can be a piezo via high impedance piezo input, microphone via XLR, or a low impedance synth input.
	\item speaking of synth input, it is designed to go very well with the stereo gesture output of the Sidrazzi organ 
	\item the COCO devices on this version are the most fully tricked out ever. they include a standardized "dolby" filter to reduce silence-noise, which also can have a side effect of punching silence into that which is looped. alongside the dolby filter are verso and inverso VCA inputs for both input and feedback. speed knob, sp.af node (control of clock frequency), and affect knob, that allows you to dial how drastic or subtle to modulate. "iron cross" output node, that is a sort of "universal" frequency readout of the actual digital circuitry itself. flip and skip inputs, as always. sample switch enables the device to loop any sample FOREVER as long as power is continuous.
	\item the QUANTUSSY in the middle, a pentagon design of oscillators, can demonstrate serious chaos, at low frequencies as well as audio rate... thus it is specially designed to give many options for modulating the COCO
\end{itemize}

\newpage

\subsection{Sobre Clicker}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/clicker}.

The Clicker is a dual voice synthesizer with 2 trigger inputs (accepting Ciat, standard trigger, or eurorack voltages), 2 manual trigger buttons, 1 music mode button to change between the two modes, 2 AD envelope generators, 2 ringlers (hordijk inspired ring modulator), with a cross modulation section. It excels at making incredibly short bursts of sound in click mode, and more melodic sounds in music mode.  Each side comprises of 2 CV inputs, 1 trigger input, and 3 CV outs. The common ground is the black jack top middle.  The LED distinguishes between the 2 modes.

“I want to create the sound of that moment when a string is activated—not the tone of the vibrating string itself.” These were the words of Jan St. Werner (Mouse on Mars), spoken just a few months ago during a phone conversation with Peter Blasser. “It's a device that triggers the speaker, creating a spatial response,” Jan elaborated, describing the concept for an electronic music instrument he envisioned for his artistic contribution to this year's Venice Biennale. Peter was intrigued, and it didn't take long for him to build on the idea and apply his own twist to it. 

The Clicker is a unique new Ciat-Lonbarde instrument. And it has a fairly distinct sonic palette to offer. The three buttons on the front trigger sounds defined by their temporal qualities. “It's an impulse generator that can be modulated,” Jan explains. “The idea is to unfold the spectra densely compressed into a tiny signal. You open it up, spacing out the distances between the different partials.” Describing the result defies description without invoking words like outburst, sharpness, graininess, and roughness.

Where synth-makers often aim for sustained, lush sounds and expansive forms, the development of the Clicker followed a very different path, Peter noted. Indeed, its magic unfolds in the briefest moments imaginable. At first, it might not even be obvious. But as soon as your perception adjusts to grasp the nuances in these short sound events, it becomes strikingly clear.

The Clicker is a highly interactive instrument. Do you have to press the buttons to coax sounds from it? That's only one option. It also integrates seamlessly with your Eurorack system or a 606 and pairs beautifully with any Ciat-Lonbarde instrument. You can even set up feedback patches to let it trigger itself.

Jan described this instrument as “an orchestra in a microsecond.” We think it's hard to add anything to that description, and are beyond excited to finally make it available through our site.

\newpage

\subsection{Sobre Peterlin}

Texto original extraído en mayo 2026 de \url{https://web.archive.org/web/20240917121115/https://patch-point.com/products/ciat-lonbarde-peterlin}.

Ciat-Lonbarde makes a synthesizer inspired by Rob Hordijk's Benjolin

PSU now sold seperately, due to the price increase of PSU, and low price of the instrument.

Technical Specs:


\begin{itemize}
	\item 3.5mm output
	\item 3.5mm input
	\item 8 knobs
	\item wood
	\item 10 banana jacka
	\item includes universal PSU
	\item can be powered by 9v battery
\end{itemize}


\newpage

\subsection{Sobre Golden Section}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/golden-section}.

Presenting the Golden Section: 5 modules on one wood base (Peterlin, 2 x Gold, 2 x Section)

A Peterlin accompanied by supplemental modules: a gesture controller named Gold after the gilding which composes its antenna, and a voltage sampler attenuverter called Section.

Gold is a gesture controller with 2 gesture points perfect for opening and closing the filter, or tuning the oscillator up and down.  Orange is the Output for each Gold.

Section is a voltage sampler where the Green are the 2 inputs for the voltage sampling, and the 2 Blue are the CV control. Orange is the Output for each Section.  A perfect patch to get your section rolling is any 2 Orange outputs on the Golden Section, whether they are from the Peterlin or the Gold.

\newpage

\subsection{Sobre Cafeteria\textunderscore Mono}


Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/cafeteria_mono}.

The gilding on this Batch is Silver Gold.

Theremin / filter / sampler / delay / reverb

digital processing and analog control * oscillators

\newpage

\subsection{Sobre Cafeteria\textunderscore Stereo}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/cafeteria_stereo}.

The Cafeteria_stereo it's the ultimate edition of the cafe line.

It's on its core: 


\begin{itemize}
	\item 1 Preamp Section
	\item 2 Cafe Delay/looper/reverb/DSP etc
	\item 1 Quantum
	\item 2 Tierhorns Filters
\end{itemize}

By the words of Peter B:

Advertising Materials

Intel inside: an esp\_cafe running open source so you can write WETF you want in weirdness, or a web browser, smartphone synth, Star Trek filter or Bluetooth streaming (ESP32-A2DP)

Intel inside: The esp\_cafe pairs an Analog Devices audio Interface for buttery twelve bit decimal Megahertz operation, with a current-mode Norton amplifier for classic analog sound.

Line Drivers: The quantum\_experiment is hybrid, using Nortons for the critical duties: VCO, VCF, and VCA, and adds quality line driver amplifiers to drive the quantization features.

Dubstep Filter: The tier\_horn gives you classic stereo enveloped sine gestures, like the deerhorn, but with a true thereminic (radio heterodyne) core.

Clara's Question: To Lev, please make me a polyphonic theremin. Sincerely, Clara.

Lev's Answer: I'm sorry Clara it's not possible; we only have two hands. Cheers, Lev.

Analog Tachometer: To answer this question, the tier\_horn creates a CV by converting the sine to a square, then pulsing an integrator, an alchemy predating phase locked loops.

Boundary pusher: To answer this question, the tier\_horn samples the boundaries of your radio gesture on the gold and lets you control how languid they decay to your rest position.

Error signal: To answer this question, The tier\_horn then references these two boundaries to create an error signal, which is like an envelope so you thus have silence on the stereo output.

Dubstep Filter: The tier\_horn uses your square waves update on “that RCA sound,” combining the boundary pusher and a secret dubstep wobbler, to “you guessed it,” filter them.

(Mister Norton is responsible for the wobbler and the filter)

Mono Stereo: if you choose country music then lets do mono, standard eurorack. If you do the new complexity, buy it in stereo.

Currency Loopback: The currency is your mini cord (stereo or mono). Connect the inputs to the outputs in a crosswise fashion.

Live Clock: The voltage control for the live clock goes directly into a current converter and analog oscillator, which the esp32 evaluates just-in-time for the livest control.

Positivity Boost: The tier\_horn attenu-verter, a knob with the purple banana, gives you full inversion, but also a boost on the version. Twice the positivity and half the negativity!

Use your ears: Patch any orange point to a green banana and twiddle the knobs.

\newpage

\subsection{Sobre Cafe Quantum}

Texto original extraído en mayo 2026 de \url{https://patch-point.com/products/cafe-quantum}.

Now building.  Shipping 2 - 4 weeks 

These are of course entirely gilded and assembled by hand, so we do of course ship in the order of payment received/order placed. 

1 Preamp Section

2 Cafe Delay/looper/reverb/DSP etc

1 Quantum